Randomise papers

### 2023-11-27 Randomisation/Randomise_Papers.qmd
completed randomisation document

```{r}
#| message: false
library("knitr")
library("tidyverse") 
library("readxl")

@@

# From the excel review spreadsheet, read in only the identifying characteristics 
identifierSheet <- readxl::read_xlsx("Spreadsheets/Stage1_Spreadsheet.xlsx", 
                                     sheet = "Metadata", skip = 1) %>%
  # Drop rows prior to start of data
  slice(4:nrow(.)) %>%
  # Only need google search query and the ID reference
  select(Review_ID, Query = `Search terms`)

# Repeat with the study metadata
studyDataSheet <- readxl::read_xlsx("Spreadsheets/Stage1_Spreadsheet.xlsx", 
                                    sheet = "Study data", skip = 1) %>%
  # Cut out non-data rows
  slice(4:nrow(.)) %>%
  # Only need year data and ID reference
  select(Review_ID, Year)

# Merge paper queries and metadata
paperList <- full_join(identifierSheet, studyDataSheet, 
                       by = "Review_ID")

# Exclude some non-query papers and add year periods to the rest
includedPapers <- paperList %>%
  # Drop empty rows, excluded, and incidental papers
  filter(!is.na(Query) & Query != "INCIDENTAL" & Year != "EXCLUDE") %>%
  # Break into time periods 
  mutate(YearPeriod = case_when(Year %in% as.character(2000:2005) ~ "2000:2005",
                                Year %in% as.character(2006:2010) ~ "2006:2010",
                                Year %in% as.character(2011:2015) ~ "2011:2015",
                                Year %in% as.character(2016:2021) ~ "2016:2021"))

head(includedPapers) %>%
  kable()
```

@@

```{r Assign random numbers}
# Randomly rank papers, then rank within the strata
rankedPapers <- includedPapers %>%
  # Add random ranks for all papers
  mutate(randRank = sample(x = 1:nrow(.), size = nrow(.))) %>%
  # Re-rank within strata
  group_by(Query, YearPeriod) %>%
  arrange(Query, YearPeriod, randRank) %>%
  mutate(GroupRank = row_number())

# Identify which papers qualify to be read
highlightedPapers <- rankedPapers %>%
  group_by(Query, YearPeriod) %>%
  mutate(PercentRank =  GroupRank/max(GroupRank)) %>%
  mutate(ToRead = case_when(
    # If there are fewer than 5, read them all
    max(GroupRank) <= 5 ~ "YES",
    # If there are 20 or fewer, read the top 5
    max(GroupRank) <= 20 & GroupRank <= 5 ~ "YES",
    max(GroupRank) <= 20 & GroupRank >= 5 ~ "NO",
    # If there are 21 or more, read the top 25%
    max(GroupRank) >= 21 & PercentRank <= 0.25  ~ "YES",
    max(GroupRank) >= 21 & PercentRank > 0.25 ~ "NO")) %>%
  ungroup() %>%
  unite(col = Group, c(Query, YearPeriod)) %>%
  # Get Review_ID, GroupRank, and ToRead
  select(Review_ID, Group, OverallRank = randRank, GroupRank, ToRead)

head(highlightedPapers) %>%
  kable()
```

@@

```{r re-add unranked papers}
unrankedPapers <- paperList %>% 
  # Drop empty rows, excluded, and incidental papers
  filter(is.na(Query) | Query == "INCIDENTAL" | Year == "EXCLUDE") %>%
  mutate(ToRead = "NO", Group = "Unranked", GroupRank = "NA", OverallRank = "NA") %>%
  select(Review_ID, Group, OverallRank, GroupRank, ToRead)

# Combine ranked and unranked papers
outputList <- rbind(highlightedPapers, unrankedPapers) %>%
  mutate(Review_ID = as.numeric(Review_ID)) %>%
  arrange(Review_ID)

# Write to disk
write_csv(outputList, file = "Randomisation/PrimarySampleRanks.csv")
```

@@

>>>>

```{r load packages}
#| message: false
library("knitr")
library("tidyverse") 
library("readxl")

@@

# From the excel review spreadsheet, read in only the identifying characteristics 
first_idSheet <- readxl::read_xlsx("Spreadsheets/Stage1_Spreadsheet.xlsx", 
                                     sheet = "Metadata", skip = 1) %>%
  # Drop rows prior to start of data
  slice(4:nrow(.)) %>%
  # Only need google search query and the ID reference
  select(Review_ID, Query = `Search terms`)

# Repeat with the study metadata
first_studyDataSheet <- readxl::read_xlsx("Spreadsheets/Stage1_Spreadsheet.xlsx", 
                                    sheet = "Study data", skip = 1) %>%
  # Cut out non-data rows
  slice(4:nrow(.)) %>%
  # Only need year data and ID reference
  select(Review_ID, Title, Year)

# Merge paper queries and metadata
first_paperList <- full_join(first_idSheet, first_studyDataSheet, 
                             by = "Review_ID") %>%
  # Exclude some non-query papers and add year periods to the rest
  # Drop empty rows, excluded, and incidental papers
  filter(!is.na(Query) & Query != "INCIDENTAL" & Year != "EXCLUDE") %>%
  # Break into time periods 
  mutate(YearPeriod = case_when(Year %in% as.character(2000:2005) ~ "2000:2005",
                                Year %in% as.character(2006:2010) ~ "2006:2010",
                                Year %in% as.character(2011:2015) ~ "2011:2015",
                                Year %in% as.character(2016:2021) ~ "2016:2021"))

head(first_paperList) %>%
  kable()
```

@@

```{r Assign random numbers}
# Randomly rank papers, then rank within the strata
first_rankedPapers <- first_paperList %>%
  # Add random ranks for all papers
  mutate(randRank = sample(x = 1:nrow(.), size = nrow(.))) %>%
  # Re-rank within strata
  group_by(Query, YearPeriod) %>%
  arrange(Query, YearPeriod, randRank) %>%
  mutate(GroupRank = row_number()) %>%
# Identify which papers qualify to be read
  group_by(Query, YearPeriod) %>%
  mutate(PercentRank =  GroupRank/max(GroupRank)) %>%
  mutate(ToRead = case_when(
    # If there are fewer than 5, read them all
    max(GroupRank) <= 5 ~ "YES",
    # If there are 20 or fewer, read the top 5
    max(GroupRank) <= 20 & GroupRank <= 5 ~ "YES",
    max(GroupRank) <= 20 & GroupRank >= 5 ~ "NO",
    # If there are 21 or more, read the top 25%
    max(GroupRank) >= 21 & PercentRank <= 0.25  ~ "YES",
    max(GroupRank) >= 21 & PercentRank > 0.25 ~ "NO")) %>%
  ungroup() %>%
  unite(col = Group, c(Query, YearPeriod)) %>%
  # Get Review_ID, GroupRank, and ToRead
  select(Review_ID, Title, Group, OverallRank = randRank, GroupRank, ToRead)

head(first_rankedPapers) %>%
  kable()
```

@@

```{r re-add unranked papers}
first_unrankedPapers <- first_paperList %>% 
  # Drop empty rows, excluded, and incidental papers
  filter(is.na(Query) | Query == "INCIDENTAL" | Year == "EXCLUDE") %>%
  mutate(ToRead = "NO", Group = "Unranked", GroupRank = "NA", OverallRank = "NA") %>%
  select(Review_ID, Title, Group, OverallRank, GroupRank, ToRead)

# Combine ranked and unranked papers
firstSample <- rbind(first_rankedPapers, first_unrankedPapers) %>%
  mutate(Review_ID = as.numeric(Review_ID)) %>%
  arrange(Review_ID) %>%
  mutate(Sample = "First")

# Write to disk
write_csv(firstSample, file = "Randomisation/FirstSampleRanks.csv")
```

@@

The list of papers pulled from Web of Science is loaded in - 1184 papers are represented here. Year-windows are assigned here as in the **First Sample.**

```{r Load second sample}
#| warning: false
# Load query and extract relevant columns
second_Query <- readxl::read_xls("Queries/MacKenzie WOS/MacKenzie03.xls") %>%
  select(Review_ID, Title = `Article Title`, Year = `Publication Year`) %>%
  # Assign year period
  mutate(YearPeriod = case_when(Year %in% 2000:2005 ~ "2000:2005",
                                Year %in% 2006:2010 ~ "2006:2010",
                                Year %in% 2011:2015 ~ "2011:2015",
                                Year %in% 2016:2021 ~ "2016:2021",
                                Year %in% 2022:2023 ~ "2022:2023"))

head(second_Query) %>%
  kable()
```

Papers are assigned a random number and ranked within their groups, and top-15 papers are flagged to be read. The output is saved for future reference.

```{r randomise second sample}
secondSample <- second_Query %>%
  # Add query flag to match with the first sample
  mutate(Query = "MacKenzie Cited") %>%
  # Add random ranks for all papers
  mutate(randRank = sample(x = 1:nrow(.), size = nrow(.))) %>%
  # Re-rank within strata
  group_by(Query, YearPeriod) %>%
  arrange(Query, YearPeriod, randRank) %>%
  mutate(GroupRank = row_number()) %>% 
  mutate(ToRead = case_when(
    GroupRank <= 15 ~ "YES",
    GroupRank >15 ~ "NO")) %>%
  unite(col = Group, c(Query, YearPeriod)) %>%
  mutate(Sample = "Second") %>%
  # Get Review_ID, GroupRank, and ToRead
  select(Sample, Review_ID, Title, Group, OverallRank = randRank, GroupRank, ToRead)

write_csv(secondSample, "Randomisation/SecondSampleRanks.csv")

head(secondSample) %>%
  kable()
```

## Sample summary

A basic representation of the two sample's coverage is presented here. The data presented is from after randomisation, **not** after reviewing, and thus does not represent the final coverage of the review as some papers may be excluded.

### First sample categories

```{r plot first sample categories}
firstSample %>%
  separate(Group, c("Query", "YearWindow"), sep = "_") %>%

  ggplot() +
  geom_bar(aes(x = YearWindow, fill = ToRead)) +
  scale_fill_manual(values = c("gray70", "dodgerblue2")) +
  facet_wrap(vars(Query), nrow = 2) +
  labs(title = "Coverage of the Primary Sample",
       y = "Number of Papers",
       x = "Year Window",
       fill = "Included")
```

### Second sample categories

```{r plot second sample categories}
secondSample %>%
  separate(Group, c("Query", "YearWindow"), sep = "_") %>%

  ggplot() +
  geom_bar(aes(x = YearWindow, fill = ToRead)) +
  scale_fill_manual(values = c("gray70", "firebrick3")) +
  labs(title = "Coverage of the Secondary Sample",
       y = "Number of Papers",
       x = "Year Window",
       fill = "Included")
```

### Combined sample

Adding the second sample improves coverage across the review period, particularly in the first two year-windows where fewer papers were included in the first sample.

```{r combined sample}
rbind(firstSample, secondSample) %>%
  separate(Group, c("Query", "YearWindow"), sep = "_") %>%
  filter(ToRead == "YES") %>%

  ggplot() +
  geom_bar(aes(x = YearWindow, fill = Sample)) +
  scale_fill_manual(values = c("dodgerblue2", "firebrick3")) +
  labs(title = "Coverage of combined sample",
       y = "Number of papers",
       x = "Year-window")
```
